---
title: 'Student list business'
mode: selfcontained
ext_widgets : {rCharts: [libraries/shiny]}
framework: revealjs
revealjs:
  theme: 'custom'
  transition: 'slide'
  center: 'false'
#bibliography: '../assets/bib/student_list_policy.bib'
#bibliography: '../assets/bib/eepa_student_list.bib'
#csl: '../assets/bib/apa.csl'
bibliography: '../../public_requests_eda/text/bib/eepa_student_list.bib'
csl: '../../public_requests_eda/text/bib/apa.csl'
--- #title


--- .section

# Results

--- .subsection

# RQ1


--- .subsubsection

# Broad patterns
## Filters used in order purchases

```{r orders-filters, fig.height = 5}
orders_filters %>% 
  add_row(univ_label = 'Research', filters = 'academic', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'academic', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'Research', filters = 'geographic', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'geographic', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'Research', filters = 'demographic', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'demographic', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'Research', filters = 'student preferences', is_asu = 'all', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'student preferences', is_asu = 'all', num = 0) %>% 
  mutate(
    filters_label = recode_factor(
      filters,
      'citizenship' = 'Citizenship',
      'rotc' = 'ROTC',
      'financial_aid' = 'Financial aid',
      'national_recognition_programs' = 'NRP',
      'edu_aspirations' = 'Education aspirations',
      'college_setting' = 'College setting',
      'college_studentbody' = 'College student body',
      'college_living_plans' = 'College living plans',
      'college_location' = 'College location',
      'college_type' = 'College type',
      'major' = 'Major',
      'college_size' = 'College size',
      'student preferences' = '   ',
      'first_gen_parent' = 'First generation',
      'low_ses' = 'Low SES',
      'gender' = 'Gender',
      'race' = 'Race',
      'demographic' = '  ',
      'proximity_search' = 'Proximity search',
      'county' = 'County',
      'intl' = 'International',
      'cbsa' = 'CBSA',
      'segment' = 'Segment',
      'geomarket' = 'Geomarket',
      'zip' = 'Zip code',
      'states_fil' = 'State',
      'geographic' = ' ',
      'hs_math' = 'HS math',
      'sat_reading_writing' = 'SAT reading/writing',
      'sat_reading' = 'SAT reading',
      'sat_writing' = 'SAT writing',
      'sat_math' = 'SAT math',
      'ap_score' = 'AP score',
      'rank' = 'Rank',
      'psat' = 'PSAT',
      'sat' = 'SAT',
      'gpa' = 'GPA',
      'academic' = '',
      'hsgrad_class' = 'HS grad class'
    )
  ) %>% 
  filter(!filters %in% c('hs_math', 'proximity_search', 'rotc', 'citizenship')) %>% 
  ggplot(aes(x = filters_label, y = num, fill = is_asu)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = if_else(filters %in% c('academic', 'geographic', 'demographic', 'student preferences') & univ_label == 'Research', str_c(str_to_sentence(filters), ' filters'), ''), y = 0), hjust = 0, vjust = 0.8, size = 2, fontface = 'bold') +
  geom_text(aes(y = num_total, label = if_else(!filters %in% c('academic', 'geographic', 'demographic', 'student preferences') & is_asu != 'asu', str_c(round(pct * 100), '%'), '')), hjust = -0.1, size = 2) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1))) +
  scale_fill_manual(values = rev(color_palette[c(1, 2)]), name = 'ASU') +
  xlab('') + ylab('Number of orders') +
  guides(fill = guide_legend(reverse = T)) +
  facet_wrap(~ factor(univ_label, levels = c('Research', 'MA/doctoral'))) +
  coord_flip() + 
  theme(
    legend.position = 'none'
  )
```


--- .subsubsection

# Academic filters
## GPA filter used

```{r orders-gpa, fig.height = 3}
orders_gpa_total <- orders_gpa %>% 
  group_by(univ_type) %>% 
  summarise(
    num_orders = sum(n_low)
  ) %>% 
  pull(num_orders, univ_type)

orders_gpa %>% 
  mutate(
    univ_label = recode_factor(
      univ_type,
      'research' = str_c('Research (N=', orders_gpa_total[['research']], ')'),
      'regional' = str_c('MA/doctoral (N=', orders_gpa_total[['regional']], ')')
    )
  ) %>% 
  ggplot(aes(x = factor(gpa_low, levels = c('A+', 'A', 'A-', 'B+', 'B', 'B-', 'C+', 'C', 'C-')), y = pct_low, fill = univ_label)) +
  geom_bar(position = 'dodge', stat = 'identity') +
  geom_text(aes(label = n_low), hjust = 0.5, vjust = 0, size = 2, position = position_dodge(0.9)) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('% of orders in university type')
```


---

## SAT filter used


```{r orders-sat, fig.height = 4}
orders_sat_total <- orders_sat %>% 
  filter(test_range == 'sat_min') %>% 
  group_by(univ_type) %>% 
  summarise(
    num_orders = sum(num)
  ) %>% 
  pull(num_orders, univ_type)

orders_sat %>% 
  mutate(
    range_label = recode_factor(
      test_range,
      'sat_min' = 'Minimum',
      'sat_max' = 'Maximum'
    ),
    univ_label = recode_factor(
      univ_type,
      'research' = str_c('Research (N=', orders_sat_total[['research']], ')'),
      'regional' = str_c('MA/doctoral (N=', orders_sat_total[['regional']], ')')
    )
  ) %>% 
  ggplot(aes(x = brks, y = pct, fill = univ_label)) +
  geom_bar(position = position_dodge2(reverse = T, padding = 0), stat = 'identity') +
  geom_text(aes(label = num), hjust = -0.1, size = 2, position = position_dodge2(0.9, reverse = T, padding = 0)) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1)), labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('% of orders in university type') +
  facet_wrap(~ range_label) +
  coord_flip()
```


---

## PSAT filter used

```{r orders-psat, fig.height = 4}
orders_psat_total <- orders_psat %>% 
  filter(test_range == 'psat_min') %>% 
  group_by(univ_type) %>% 
  summarise(
    num_orders = sum(num)
  ) %>% 
  pull(num_orders, univ_type)

orders_psat %>% 
  mutate(
    range_label = recode_factor(
      test_range,
      'psat_min' = 'Minimum',
      'psat_max' = 'Maximum'
    ),
    univ_label = recode_factor(
      univ_type,
      'research' = str_c('Research (N=', orders_psat_total[['research']], ')'),
      'regional' = str_c('MA/doctoral (N=', orders_psat_total[['regional']], ')')
    )
  ) %>% 
  ggplot(aes(x = brks, y = pct, fill = univ_label)) +
  geom_bar(position = position_dodge2(reverse = T, padding = 0), stat = 'identity') +
  geom_text(aes(label = num), hjust = -0.1, size = 2, position = position_dodge2(0.9, reverse = T, padding = 0)) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1)), labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('% of orders in university type') +
  facet_wrap(~ range_label) +
  coord_flip()
```



--- .subsubsection

#  Geographic filters
## State filter used by research universities, out-of-state

```{r orders-state-research-outofstate, fig.height = 2.8, fig.width = 5}
# Out-of-state
plot_usmap(
  regions = 'states',
  data = orders_state_research %>% filter(locale == 'outofstate'),
  values = 'frequency'
) +
  scale_fill_gradient2(
    low = 'white',
    mid = '#ededed',
    high = color_palette[[1]], 
    midpoint = 35,
    name = 'Number of orders',
    label = label_number(accuracy = 1),
    limits = c(0, 120)
  ) + 
  # labs(title = 'State filter used by research universities (out-of-state)') +
  theme(
    text = element_text(size = 7), 
    panel.background = element_rect(fill = '#f0f0f0', color = NA),
    plot.background = element_rect(fill = '#f0f0f0', color = NA),
    plot.title = element_text(color = '#444444', hjust = 0.5, face = 'bold'), 
    legend.background = element_rect(fill = '#f0f0f0', color = NA),
    legend.title = element_text(face = 'bold'),
    legend.position = 'right'
  )
```

---

## State filter used by research universities, in-state

```{r orders-state-research-instate, fig.height = 2.8, fig.width = 5}
# In-state
plot_usmap(
  regions = 'states',
  data = orders_state_research %>% filter(locale == 'instate'),
  values = 'frequency'
) +
  scale_fill_continuous(
    low = '#ededed',
    high = color_palette[[1]], 
    name = 'Number of orders',
    label = label_number(accuracy = 1),
    limits = c(0, 10)
  ) + 
  # labs(title = 'State filter used by research universities (in-state)') +
  theme(
    text = element_text(size = 7), 
    panel.background = element_rect(fill = '#f0f0f0', color = NA),
    plot.background = element_rect(fill = '#f0f0f0', color = NA),
    plot.title = element_text(color = '#444444', hjust = 0.5, face = 'bold'), 
    legend.background = element_rect(fill = '#f0f0f0', color = NA),
    legend.title = element_text(face = 'bold'),
    legend.position = 'right'
  )
```


--- .subsubsection

# Demographic filters
## Race filter

```{r orders-race, fig.height = 3}
orders_race %>% 
  ggplot(aes(fill = univ_label, y = count, x = reorder(race_ethnicity_group, count))) + 
  geom_bar(position = 'stack', stat = 'identity') +
  geom_text(aes(label = count), hjust = -0.1, size = 2) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('Number of orders') +
  coord_flip()
```


--- .subsubsection

# Combination of filters
## Filter combos used in order purchases

```{r orders-filters-combo, results = 'asis'}
bind_cols(
  orders_filters_combo %>% 
    filter(univ_type == 'research') %>% 
    select(filter_combos, n, pct) %>% 
    head(10) %>%
    mutate(pct = str_c(round(pct * 100), '%')),
  orders_filters_combo %>% 
    filter(univ_type == 'regional') %>% 
    select(filter_combos, n, pct) %>% 
    head(10) %>%
    mutate(pct = str_c(round(pct * 100), '%'))
) %>% 
  xtable(align = c('c', 'p', 'c', 'c', 'p', 'c', 'c'), display = c('d', 's', 'd', 's', 's', 'd', 's')) %>% 
  print(add.to.row = list(pos = list(0), command = c('<tr><th colspan="3" style="text-align:center;">Research</th><th colspan="3" style="text-align:center;">MA/doctoral</th></tr><tr><th>Filters</th><th>Count</th><th>Percent</th><th>Filters</th><th>Count</th><th>Percent</th></tr>')), include.colnames = F)
```


--- .subsection

# RQ2


---

# Characteristics of Prospects
## Number of prospects by university type and location

```{r rq2-counts, fig.height = 3}
rq2_counts %>% 
  mutate(
    region = recode_factor(
      region,
      'international' = 'International',
      'domestic' = 'Domesetic'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'regional_outofstate' = 'MA/doctoral (out-of-state)',
      'regional_instate' = 'MA/doctoral (in-state)',
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = n, fill = region)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(univ_type == 'regional' & locale == 'outofstate', '', if_else(n < 1000000, str_c(round(n / 1000), 'K'), str_c(round(n / 1000000, 1), 'M')))), size = 2, position = position_stack(vjust = 0.5)) +
  geom_text(mapping = aes(label = if_else(univ_type == 'regional' & region == 'Domesetic' & locale == 'outofstate', str_c(round(n / 1000), 'K (300 international)'), ''), x = 30e3), hjust = 0, size = 2) +
  scale_fill_manual(values = rev(color_palette[1:2]), name = '') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), labels = function(n) if_else(n < 1000000, str_c(n / 1000, 'K'), str_c(n / 1000000, 'M'))) +
  xlab('Number of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```



--- .subsubsection

# Public research universities 
## Racial composition of prospects in lists purchased

```{r rq2-race-research, fig.height = 2}
rq2_race %>% 
  filter(univ_type == 'research', region != 'international') %>% 
  mutate(
    stu_race_cb = if_else(is.na(stu_race_cb), 999, unclass(stu_race_cb)),
    race = recode_factor(
      stu_race_cb,
      `999` = 'Missing',
      `0` = 'No response',
      `10` = 'Other',
      `12` = 'Multiracial',
      `8` = 'NH/PI',
      `1` = 'AI/AN',
      `4` = 'Latinx',
      `3` = 'Black',
      `2` = 'Asian',
      `9` = 'White'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  filter(race != 'Missing') %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_fill_manual(values = rev(color_palette), name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


---

## Median household income of prospects in lists purchased

```{r rq2-income-research, fig.height = 1.5}
rq2_income %>% 
  filter(univ_type == 'research', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = color_palette[[1]]) + 
  geom_text(mapping = aes(label = str_c('$', round(income_2564 / 1000), 'K')), size = 2, hjust = -0.1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 115000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


---

## Locale of prospects in lists purchased

```{r rq2-locale-research, fig.height = 2}
rq2_locale %>% 
  filter(univ_type == 'research', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = locale_text)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Locale') +
  scale_fill_manual(values = rev(color_palette[c(1:5, 7)]), name = 'Locale') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


--- .subsubsection

# Public ma/doctoral universities
## Racial composition of prospects in lists purchased

```{r rq2-race-regional, fig.height = 2}
rq2_race %>% 
  filter(locale == 'instate', region != 'international') %>% 
  mutate(
    stu_race_cb = if_else(is.na(stu_race_cb), 999, unclass(stu_race_cb)),
    race = recode_factor(
      stu_race_cb,
      `999` = 'Missing',
      `0` = 'No response',
      `10` = 'Other',
      `12` = 'Multiracial',
      `8` = 'NH/PI',
      `1` = 'AI/AN',
      `4` = 'Latinx',
      `3` = 'Black',
      `2` = 'Asian',
      `9` = 'White'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_instate' = 'Research (in-state)',
      'regional_instate' = 'MA/doctoral (in-state)'
    )
  ) %>% 
  filter(race != 'Missing') %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_fill_manual(values = rev(color_palette), name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


---

## Median household income of prospects purchased

```{r rq2-income-regional, fig.height = 2}
rq2_income %>% 
  filter(locale == 'instate', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_instate' = 'Research (in-state)',
      'regional_instate' = 'MA/doctoral (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = color_palette[[1]]) + 
  geom_text(mapping = aes(label = str_c('$', round(income_2564 / 1000), 'K')), size = 2, hjust = -0.1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 105000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


---

## Locale of prospects in lists purchased

```{r rq2-locale-regional, fig.height = 2}
rq2_locale %>% 
  filter(locale == 'instate', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_instate' = 'Research (in-state)',
      'regional_instate' = 'MA/doctoral (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = locale_text)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Locale') +
  scale_fill_manual(values = rev(color_palette[c(1:5, 7)]), name = 'Locale') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


--- .subsection

# RQ3


--- .subsubsection

# Characteristics by filters
## Prospect characteristics across individual filter criteria

```{r rq3, results = 'asis'}
rq3 %>% 
  mutate_if(is.numeric, ~if_else(. < 1, . * 100, .)) %>% 
  mutate_if(is.numeric, round, 0) %>% 
  mutate_if(is.numeric, ~if_else(row_number() == 18, str_c('$', round(. / 1000), 'K'), prettyNum(., big.mark = ','))) %>% 
  mutate(
    row_subj = recode_factor(
      row_subj,
      'n' = ' ',
      'pct_instate' = '% In-state',
      'pct_outofstate' = '% Out-of-state',
      'pct_white' = '% White',
      'pct_asian' = '% Asian',
      'pct_black' = '% Black',
      'pct_latinx' = '% Latinx',
      'pct_aian' = '% AI/AN',
      'pct_nhpi' = '% NH/PI',
      'pct_multiracial' = '% Multiracial',
      'pct_raceother' = '% Other',
      'pct_noresponse' = '% No response',
      'pct_racemissing' = '% Missing',
      'pct_male' = '% Male',
      'pct_female' = '% Female',
      'pct_genderother' = '% Other ',
      'pct_gendermissing' = '% Missing ',
      'med_income' = 'Median income',
      'pct_city' = '% City',
      'pct_suburban' = '% Suburban',
      'pct_fringe' = '% Rural - Fringe',
      'pct_distant' = '% Rural - Distant',
      'pct_remote' = '% Rural - Remote',
      'pct_localemissing' = '% Missing  '
    )
  ) %>%
  arrange(row_subj) %>% 
  add_column('x1' = if_else(.$row_subj == ' ', 'Total', ' '), .before = 'row_subj') %>% 
  add_column('x2' = ' ', .before = 'filter_zip') %>% 
  add_column('x3' = ' ', .before = 'filter_race') %>% 
  xtable(align = c('c', 'l', 'l', rep('c', 15))) %>% 
  print(
    html.table.attributes = 'style="font-size:10px;"',
    add.to.row = list(
      pos = list(0, 1, 3, 13, 17, 18),
      command = c('<tr style="text-align: center;"><th colspan="3" style="border-bottom: none;"></th><th colspan="5">Academic</th><th style="border-bottom: none;"><th colspan="5">Geographic</th><th style="border-bottom: none;"></th><th colspan="2">Demographic</th></tr><tr style="text-align:center;"><th width="15px;"></th><th></th><th>All domestic</th><th>GPA</th><th>PSAT</th><th>SAT</th><th>HS rank</th><th>AP score</th><th width="5px;"></th><th>Zip code</th><th>State</th><th>Geomarket</th><th>Segment</th><th>CBSA</th><th width="5px;"></th><th>Race</th><th>Gender</th></tr>', '<tr style="font-weight:900"><td colspan="2">Location</td></tr>', '<tr style="font-weight:900"><td colspan="2">Race/ethnicity</td></tr>', '<tr style="font-weight:900"><td colspan="2">Gender</td></tr>', '<tr style="font-weight:900"><td colspan="2">Household income</td></tr>', '<tr style="font-weight:900"><td colspan="2">Locale</td></tr>')
    ),
    include.colnames = F
  )
```


--- .subsubsection

# Zip code & test score filters
## Los Angeles prospects from top income decile zip codes

```{r asu-la-deep-dive, fig.height = 3}
asu_la_figure_totals <- asu_la_full %>% 
  mutate(
    count_type = recode_factor(
      in_zip_top10pct,
      `0` = 'Rest',
      `1` = 'Top 10%'
    ),
    ord_type = recode_factor(
      ord_type,
      'psat_high' = 'High PSAT order (1270-1520)',
      'psat_med' = 'Medium PSAT order (1190-1260)',
      'psat_low' = 'Low PSAT order (1110-1210)',
      'sat_med' = 'Medium SAT order (1140-1260)'
    )
  ) %>% 
  group_by(ord_type, count_type) %>% 
  summarise(
    count = sum(count)
  ) %>% 
  ungroup()

asu_la %>% 
  mutate(
    stu_race_cb = if_else(is.na(stu_race_cb), 999, unclass(stu_race_cb)),
    race = recode_factor(
      stu_race_cb,
      `999` = 'Missing',
      `0` = 'No response',
      `10` = 'Other',
      `12` = 'Multiracial',
      `8` = 'NH/PI',
      `1` = 'AI/AN',
      `4` = 'Latinx',
      `3` = 'Black',
      `2` = 'Asian',
      `9` = 'White'
    ),
    count_type = recode_factor(
      in_zip_top10pct,
      `0` = 'Rest',
      `1` = 'Top 10%'
    ),
    ord_type = recode_factor(
      ord_type,
      'psat_high' = 'High PSAT order (1270-1520)',
      'psat_med' = 'Medium PSAT order (1190-1260)',
      'psat_low' = 'Low PSAT order (1110-1210)',
      'sat_med' = 'Medium SAT order (1140-1260)'
    )
  ) %>% 
  filter(!race %in% c('Missing', 'No response')) %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  geom_text(data = asu_la_figure_totals, mapping = aes(x = 1, fill = NULL, label = str_c('N=', prettyNum(count, ','))), size = 2, hjust = -0.1) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_fill_manual(values = rev(color_palette[1:7]), name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 1.15)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  facet_wrap(~ ord_type) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1)
  )
```



--- .subsubsection

# Geodemographic segment filters
## Filter by neighborhood segments

```{r orders-cluster-en, results = 'asis'}
en_df <- data.frame(
  cluster = c(51:83, 'Total'),
  sat_math = c(546, 480, 561, 458, 566, 420, 541, 533, 561, 589, 585, 596, 548, 466, 440, 499, 519, 552, 534, 613, 405, 399, 528, 433, 459, 514, 502, 594, 550, 534, 491, 496, 500, 512),
  sat_cr = c(533, 470, 544, 443, 565, 411, 519, 489, 562, 590, 567, 595, 541, 466, 433, 492, 501, 558, 521, 598, 408, 397, 514, 435, 457, 509, 492, 578, 551, 527, 483, 491, 490, 502),
  pct_outofstate = c(0.32, 0.3, 0.32, 0.25, 0.52, 0.29, 0.52, 0.28, 0.52, 0.63, 0.51, 0.67, 0.39, 0.48, 0.23, 0.2, 0.27, 0.52, 0.37, 0.65, 0.39, 0.31, 0.29, 0.29, 0.28, 0.27, 0.26, 0.56, 0.57, 0.39, 0.27, 0.29, 0.19, 0.32),
  pct_nonwhite = c(0.3, 0.58, 0.5, 0.83, 0.24, 0.93, 0.47, 0.87, 0.24, 0.37, 0.3, 0.24, 0.23, 0.34, 0.93, 0.12, 0.53, 0.35, 0.19, 0.29, 0.97, 0.87, 0.42, 0.84, 0.85, 0.38, 0.18, 0.26, 0.32, 0.39, 0.57, 0.21, 0.26, 0.43),
  pct_aid = c(0.57, 0.71, 0.55, 0.76, 0.63, 0.66, 0.43, 0.69, 0.74, 0.36, 0.4, 0.72, 0.65, 0.29, 0.78, 0.76, 0.59, 0.65, 0.65, 0.61, 0.68, 0.47, 0.62, 0.79, 0.72, 0.64, 0.75, 0.39, 0.74, 0.65, 0.72, 0.75, 0.71, 0.65),
  med_income = c(95432, 63578, 92581, 38977, 71576, 35308, 67394, 68213, 54750, 104174, 123858, 59824, 69347, 49829, 45081, 50453, 60960, 57902, 88100, 86381, 42661, 32708, 90849, 44065, 50421, 61332, 62372, 134400, 40909, 49877, 63030, 53465, 49335, 70231)
) %>% 
  mutate(
    pct_outofstate = percent(pct_outofstate, accuracy = 1),
    pct_nonwhite = percent(pct_nonwhite, accuracy = 1),
    pct_aid = percent(pct_aid, accuracy = 1),
    med_income = dollar(med_income)
  ) %>% 
  xtable(align = c('l', 'l', rep('c', 6))) %>% 
  print(type = 'html', html.table.attributes = 'style="font-size:10px;" id="cluster-en"', add.to.row = list(
      pos = list(0),
      command = c('<tr style="text-align:center;"><th style="text-align:left;">2011 D+ Cluster</th><th>SAT Math</th><th>SAT CR</th><th>Going Out of State</th><th>Percent NonWhite</th><th>Need Financial Aid</th><th>Med Income</th></tr>')),
      include.colnames = F
  )
```


---

## Filter by high school segments

```{r orders-cluster-hs, results = 'asis'}
hs_df <- data.frame(
  cluster = c(51:79, 'Total'),
  sat_math = c(462, 489, 471, 376, 489, 536, 434, 592, 499, 523, 485, 474, 440, 606, 515, 498, 526, 541, 390, 595, 400, 528, 451, 654, 514, 600, 595, 473, 594, 514),
  sat_cr = c(457, 496, 484, 371, 481, 508, 435, 577, 489, 549, 370, 473, 427, 542, 503, 515, 546, 540, 395, 581, 412, 544, 438, 579, 502, 584, 508, 468, 585, 502),
  pct_outofstate = c(0.14, 0.81, 0.28, 0.33, 0.39, 0.73, 0.29, 0.51, 0.19, 0.23, 0.33, 0.34, 0.28, 0.37, 0.28, 0.37, 0.48, 0.41, 0.36, 0.56, 0.57, 0.35, 0.24, 0.76, 0.31, 0.72, 0.64, 0.48, 0.61, 0.32),
  pct_nonwhite = c(0.33, 0.99, 0.38, 0.96, 0.46, 0.43, 0.82, 0.27, 0.18, 0.3, 0.89, 0.92, 0.86, 0.89, 0.43, 0.37, 0.41, 0.26, 0.92, 0.33, 0.98, 0.25, 0.89, 0.8, 0.2, 0.5, 0.75, 0.43, 0.26, 0.44),
  pct_aid = c(0.68, 0.77, 0.62, 0.38, 0.44, 0.49, 0.79, 0.32, 0.74, 0.33, 0.09, 0.67, 0.72, 0.57, 0.65, 0.73, 0.69, 0.62, 0.74, 0.48, 0.8, 0.64, 0.76, 0.46, 0.71, 0.28, 0.39, 0.22, 0.71, 0.65),
  med_income = c(40918, 64730, 60833, 38146, 71845, 63967, 48301, 104509, 47685, 70175, 61385, 55515, 49238, 81911, 72692, 60272, 71279, 79260, 43391, 105721, 43137, 70018, 48406, 59089, 72850, 90265, 39490, 56703, 65180, 70223)
) %>% 
  mutate(
    pct_outofstate = percent(pct_outofstate, accuracy = 1),
    pct_nonwhite = percent(pct_nonwhite, accuracy = 1),
    pct_aid = percent(pct_aid, accuracy = 1),
    med_income = dollar(med_income)
  ) %>% 
  xtable(align = c('l', 'l', rep('c', 6))) %>% 
  print(type = 'html', html.table.attributes = 'style="font-size:10px;" id="cluster-hs"', add.to.row = list(
      pos = list(0),
      command = c('<tr style="text-align:center;"><th style="text-align:left;">2011 D+ Cluster</th><th>SAT Math</th><th>SAT CR</th><th>Going Out of State</th><th>Percent NonWhite</th><th>Need Financial Aid</th><th>Med Income</th></tr>')),
      include.colnames = F
  )
```


---

## Segment filter prospects by metro

```{r uiuc-deep-dive, fig.height = 3.6}
uiuc_income_figure <- uiuc_income %>% 
  mutate(
    count_type = recode_factor(
      ord_type,
      'prospect' = 'Prospects',
      'metro' = 'Metro'
    ),
    cbsa_name = str_replace(cbsa_name, ' Metro Area', '')
  ) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = extra_colors[[1]]) + 
  geom_text(mapping = aes(label = str_c('$', round(income_2564 / 1000), 'K')), size = 2, hjust = -0.1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 250000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA', 'Los Angeles-Long Beach-Anaheim, CA', 'Philadelphia-Camden-Wilmington, PA-NJ-DE-MD', 'Washington-Arlington-Alexandria, DC-VA-MD-WV')), ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm'))
  )

uiuc_race_figure_data <- uiuc_race %>% 
  mutate(
    race = recode_factor(
      race,
      'unknown' = 'Missing',
      'noresponse' = 'No response',
      'other' = 'Other',
      'tworaces' = 'Multiracial',
      'nativehawaii' = 'NH/PI',
      'amerindian' = 'AI/AN',
      'hispanic' = 'Latinx',
      'black' = 'Black',
      'asian' = 'Asian',
      'white' = 'White'
    ),
    count_type = recode_factor(
      ord_type,
      'prospect' = 'Prospects',
      'metro' = 'Metro'
    ),
    cbsa_name = str_replace(cbsa_name, ' Metro Area', '')
  ) 

uiuc_race_figure_totals <- uiuc_race_full %>% 
  mutate(
    count_type = recode_factor(
        ord_type,
        'prospect' = 'Prospects',
        'metro' = 'Metro'
    )
  ) %>% 
  group_by(cbsa_name, count_type) %>% 
  summarise(
    count = sum(count)
  ) %>% 
  ungroup()
  
uiuc_race_figure <- uiuc_race_figure_data %>% 
  filter(!race %in% c('Missing', 'No response')) %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  geom_text(data = uiuc_race_figure_totals, mapping = aes(x = 1, fill = NULL, label = if_else(count_type != 'Metro', str_c('N=', prettyNum(count, big.mark = ',')), '')), size = 2, hjust = -0.15) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_fill_manual(values = rev(color_palette[1:7]), name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 1.14)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA', 'Los Angeles-Long Beach-Anaheim, CA', 'Philadelphia-Camden-Wilmington, PA-NJ-DE-MD', 'Washington-Arlington-Alexandria, DC-VA-MD-WV')), ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    strip.text.x = element_text(color = NA, size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    plot.margin = unit(c(5.5, 0, 5.5, -45), 'pt')
  )

grid.arrange(
  uiuc_income_figure,
  uiuc_race_figure,
  ncol = 2
)
```


---

## Segment filter prospects interactive map

<iframe src="https://mpatricia01.github.io/public_requests_eda/outputs/maps/map_segment_green.html" id="uiuc-deep-dive-map" width=100% height=100% allowtransparency="true"></iframe>


--- .subsubsection

# Women in STEM
## Women in STEM prospects by metro

```{r ucsd-deep-dive, fig.height = 4}
ucsd_income_figure <- ucsd_income %>% 
  mutate(
    count_type = recode_factor(
      ord_type,
      'sat' = 'SAT',
      'ap' = 'AP',
      'metro' = 'Metro'
    ),
    cbsa_name = str_replace(str_replace(cbsa_name, 'Roswell', 'Alpharetta'), ' Metro Area', '')
  ) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = extra_colors[[1]]) + 
  geom_text(mapping = aes(label = str_c('$', round(income_2564 / 1000), 'K')), size = 2, hjust = -0.1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 163000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ cbsa_name, ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm'))
  )

ucsd_race_figure_data <- ucsd_race %>% 
  mutate(
    race = recode_factor(
      race,
      'missing' = 'Missing',
      'noresponse' = 'No response',
      'other' = 'Other',
      'tworaces' = 'Multiracial',
      'nativehawaii' = 'NH/PI',
      'amerindian' = 'AI/AN',
      'hispanic' = 'Latinx',
      'black' = 'Black',
      'asian' = 'Asian',
      'white' = 'White'
    ),
    count_type = recode_factor(
      ord_type,
      'sat' = 'SAT',
      'ap' = 'AP',
      'metro' = 'Metro'
    ),
    cbsa_name = str_replace(cbsa_name, ' Metro Area', '')
  )

ucsd_race_figure_totals <- ucsd_race_full %>% 
  mutate(
    count_type = recode_factor(
      ord_type,
      'sat' = 'SAT',
      'ap' = 'AP',
      'metro' = 'Metro'
    )
  ) %>% 
  group_by(cbsa_name, count_type) %>% 
  summarise(
    count = sum(count)
  ) %>% 
  ungroup()
  
ucsd_race_figure <- ucsd_race_figure_data %>% 
  filter(!race %in% c('Missing', 'No response')) %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  geom_text(data = ucsd_race_figure_totals, mapping = aes(x = 1, fill = NULL, label = if_else(count_type != 'Metro', str_c('N=', count), '')), size = 2, hjust = -0.2) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_fill_manual(values = rev(color_palette[1:7]), name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 1.14)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ cbsa_name, ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    strip.text.x = element_text(color = NA, size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    plot.margin = unit(c(5.5, 0, 5.5, 2), 'pt')
  )

grid.arrange(
  ucsd_income_figure,
  ucsd_race_figure,
  ncol = 2,
  widths = c(2, 3)
)
```


--- .subsubsection

# Targeting URM students
## Race and ethnicity variables, aggregated vs. alone

```{r poc-race-deep-dive, fig.height = 4.5}
poc_cb_totals <- poc_cb %>%
  group_by(cbsa_code) %>% 
  summarise(total = sum(count)) %>% 
  ungroup()

poc_cb_figure <- poc_cb %>% 
  left_join(poc_cb_totals, by = 'cbsa_code') %>% 
  mutate(
    race = recode_factor(
      race,
      'unknown' = 'Missing',
      'noresponse' = 'No response',
      'other' = 'Other',
      'tworaces' = 'Multiracial',
      'nativehawaii' = 'NH/PI',
      'amerindian' = 'AI/AN',
      'black' = 'Black',
      'asian' = 'Asian',
      'white' = 'White',
      'hispanic' = 'Latinx'
    ),
    cbsa_name = str_c(cbsa_name, ' (N=', total, ')')
  ) %>% 
  ggplot(aes(x = race, y = pct)) +
  geom_bar(stat = 'identity', fill = color_palette[[1]]) +
  geom_text(aes(label = str_c(round(pct * 100), '%')), hjust = -0.1, size = 2) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1)), limits = c(0, 1)) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA (N=949)', 'Miami-Fort Lauderdale-West Palm Beach, FL (N=671)', 'Houston-The Woodlands-Sugar Land, TX (N=371)')), ncol = 1) +
  coord_flip() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    plot.margin = unit(c(5.5, 0, 5.5, 0), 'pt')
  )

poc_common_figure <- poc_common %>% 
  mutate(
    race = recode_factor(
      race,
      'native_hawaiian' = 'NH/PI',
      'american_indian' = 'AI/AN',
      'black' = 'Black',
      'asian' = 'Asian',
      'white' = 'White',
      'is_hisp' = 'Latinx'
    )
  ) %>% 
  ggplot(aes(x = race, y = pct)) +
  geom_bar(stat = 'identity', fill = color_palette[[1]]) +
  geom_text(aes(label = str_c(round(pct * 100), '%')), hjust = -0.1, size = 2) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA', 'Miami-Fort Lauderdale-West Palm Beach, FL', 'Houston-The Woodlands-Sugar Land, TX')), ncol = 1) +
  coord_flip() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(color = NA, size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    strip.background = element_rect(fill = '#f0f0f0', color = NA),
    plot.margin = unit(c(5.5, 0, 5.5, -20), 'pt')
  )

poc_cb_figure + poc_common_figure + plot_annotation(title = 'Aggregate race/ethnicity variable                                                                        Multiple races and ethnicities', theme = theme(plot.title = element_text(size = 7, color = 'black')))
```


---

## Purchased profiles for students of color by metro

```{r poc-prospects-deep-dive, fig.height = 4.5}
poc_hs_figure <- poc_hs %>% 
  left_join(poc_cb_totals, by = 'cbsa_code') %>% 
  mutate(
    control = recode_factor(
      control,
      'public' = 'Public HS',
      'private' = 'Private HS'
    ),
    count_type = recode_factor(
      ord_type,
      'prospect' = 'Prospects',
      'metro' = 'Metro'
    ),
    cbsa_name = str_c(cbsa_name, ' (N=', total, ')')
  ) %>% 
  filter(control == 'Private HS') %>% 
  ggplot(aes(y = count_type, x = pct, fill = control)) + 
  geom_bar(stat = 'identity', fill = extra_colors[[2]]) + 
  geom_text(mapping = aes(label = str_c(round(pct * 100), '% ', control)), size = 2, hjust = -0.05) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 1.5)) +
  # scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA (N=949)', 'Miami-Fort Lauderdale-West Palm Beach, FL (N=671)', 'Houston-The Woodlands-Sugar Land, TX (N=371)')), ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    plot.margin = unit(c(5.5, 0, 5.5, -10), 'pt')
  )

poc_income_figure <- poc_income %>% 
  add_row(cbsa_code = '26420', cbsa_name = 'Houston-The Woodlands-Sugar Land, TX', ord_type = 'public', income_2564 = 0) %>% 
  add_row(cbsa_code = '26420', cbsa_name = 'Houston-The Woodlands-Sugar Land, TX', ord_type = 'private', income_2564 = 0) %>% 
  add_row(cbsa_code = '33100', cbsa_name = 'Miami-Fort Lauderdale-West Palm Beach, FL', ord_type = 'public', income_2564 = 0) %>% 
  add_row(cbsa_code = '33100', cbsa_name = 'Miami-Fort Lauderdale-West Palm Beach, FL', ord_type = 'private', income_2564 = 0) %>% 
  add_row(cbsa_code = '35620', cbsa_name = 'New York-Newark-Jersey City, NY-NJ-PA', ord_type = 'public', income_2564 = 0) %>% 
  add_row(cbsa_code = '35620', cbsa_name = 'New York-Newark-Jersey City, NY-NJ-PA', ord_type = 'private', income_2564 = 0) %>% 
  mutate(
    count_type = recode_factor(
      ord_type,
      'private_1+' = '1+',
      'private_zero' = ' No buys',
      'private' = ' ',
      'public_6+' = '6+',
      'public_1-5' = '1-5',
      'public_zero' = 'No buys',
      'public' = '',
      'prospect' = 'Prospects',
      'metro' = 'Metro'
    ),
    cbsa_name = str_replace(str_replace(cbsa_name, 'Pompano', 'West Palm'), ' Metro Area', '')
  ) %>% 
  filter(str_detect(ord_type, 'private', T)) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = extra_colors[[1]]) + 
  geom_text(mapping = aes(label = if_else(income_2564 == 0, '', str_c('$', round(income_2564 / 1000), 'K'))), size = 2, hjust = -0.1) +
  geom_text(aes(label = if_else(income_2564 != 0, '', if_else(count_type == '', 'Public HS', 'Private HS')), x = 0), hjust = 0, vjust = 0.9, size = 2, fontface = 'bold') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 350000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA', 'Miami-Fort Lauderdale-West Palm Beach, FL', 'Houston-The Woodlands-Sugar Land, TX')), ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(color = NA, size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    plot.margin = unit(c(5.5, 0, 5.5, -120), 'pt')
  )

poc_race_figure <- poc_race %>% 
  # filter(count != 0) %>% 
  add_row(cbsa_code = '26420', cbsa_name = 'Houston-The Woodlands-Sugar Land, TX', ord_type = 'public', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '26420', cbsa_name = 'Houston-The Woodlands-Sugar Land, TX', ord_type = 'private', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '26420', cbsa_name = 'Houston-The Woodlands-Sugar Land, TX', ord_type = 'prospect', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '26420', cbsa_name = 'Houston-The Woodlands-Sugar Land, TX', ord_type = 'metro', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '33100', cbsa_name = 'Miami-Fort Lauderdale-West Palm Beach, FL', ord_type = 'public', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '33100', cbsa_name = 'Miami-Fort Lauderdale-West Palm Beach, FL', ord_type = 'private', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '33100', cbsa_name = 'Miami-Fort Lauderdale-West Palm Beach, FL', ord_type = 'prospect', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '33100', cbsa_name = 'Miami-Fort Lauderdale-West Palm Beach, FL', ord_type = 'metro', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '35620', cbsa_name = 'New York-Newark-Jersey City, NY-NJ-PA', ord_type = 'public', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '35620', cbsa_name = 'New York-Newark-Jersey City, NY-NJ-PA', ord_type = 'private', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '35620', cbsa_name = 'New York-Newark-Jersey City, NY-NJ-PA', ord_type = 'prospect', race = 'white', pct = 0) %>% 
  add_row(cbsa_code = '35620', cbsa_name = 'New York-Newark-Jersey City, NY-NJ-PA', ord_type = 'metro', race = 'white', pct = 0) %>% 
  mutate(
    race = recode_factor(
      race,
      'unknown' = 'Missing',
      'tworaces' = 'Multiracial',
      'nativehawaii' = 'NH/PI',
      'amerindian' = 'AI/AN',
      'hispanic' = 'Latinx',
      'black' = 'Black',
      'asian' = 'Asian',
      'white' = 'White'
    ),
    count_type = recode_factor(
      ord_type,
      'private_1+' = '1+',
      'private_zero' = ' No buys',
      'private' = ' ',
      'public_6+' = '6+',
      'public_1-5' = '1-5',
      'public_zero' = 'No buys',
      'public' = '',
      'prospect' = '  ',
      'metro' = '   '
    )
  ) %>% 
  filter(!race %in% c('Missing', 'No response'), str_detect(ord_type, 'private', T)) %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5), color = color_text) +
  geom_text(aes(label = if_else(str_detect(count_type, '\\w'), '', if_else(count_type == '', 'Public HS', '')), x = 0), hjust = 0, vjust = 0.9, size = 2, fontface = 'bold') +
  geom_text(aes(x = 1, fill = NULL, label = if_else(str_detect(count_type, '\\w'), str_c(' ', num_hs, ' HS', if_else(num_prospects > 0, str_c('\n ', num_prospects, ' prospects'), '')), '')), size = 1, hjust = 0) +
  # scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_fill_manual(values = rev(color_palette[1:7]), name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 1.14)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  facet_wrap(~ factor(cbsa_name, levels = c('New York-Newark-Jersey City, NY-NJ-PA', 'Miami-Fort Lauderdale-West Palm Beach, FL', 'Houston-The Woodlands-Sugar Land, TX')), ncol = 1) +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    # plot.background = element_rect(color = 'gray'),
    # panel.background = element_rect(color = 'red'),
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 6, face = 'bold', hjust = 1),
    strip.text.x = element_text(color = NA, size = 7, face = 'bold', hjust = 0, margin = margin(2, 0, 2, 0, 'mm')),
    legend.margin = unit(c(0, 0, 0, 0), 'pt'),
    plot.margin = unit(c(4, 0, 5.5, -100), 'pt')
  )

grid.arrange(
  poc_hs_figure,
  poc_income_figure,
  poc_race_figure,
  ncol = 3,
  widths = c(2, 1, 1)
)
```



---

## Purchased profiles for students of color interactive map

<iframe src="https://mpatricia01.github.io/public_requests_eda/outputs/maps/map_poc_green.html" id="poc-prospects-deep-dive-map" width=100% height=100% allowtransparency="true"></iframe>

