---
title: "The Student List Business"
  # "Deconstructing the student list business in higher education"
author: 
  - Ozan Jaquette
  - Karina Salazar
  - Patrica Martín
bibliography: ./../assets/bib/student_list_policy.bib # "./../../../Google Drive/student_list_brief/literature/student_list_policy.bib"
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../assets/bib/apa.csl
  # C:\Users\ozanj\Documents\student_list_policy\assets\bib
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 12pt
---

```{r, include = FALSE, eval = FALSE}
# filepath location of apa.csl file and location of .bib list of references
getwd()
list.files(path = file.path('.','..','assets','bib')) # apa.csl file
list.files(path = file.path('.','..','..','..','Google Drive','student_list_brief','literature')) # .bib reference list on Google Drive
```


```{r setup, include = FALSE}
library(knitr)
library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(scales)

knitr::opts_chunk$set(echo = F, message = F, warning = F)

knitr::knit_hooks$set(inline = function(x) {   if(!is.numeric(x)){     x   }else{    prettyNum(round(x,2), big.mark=",")    } })

theme_set(
  theme(
    text = element_text(size = 7),
    panel.background = element_blank(),
    plot.title = element_text(color = '#444444', size = 7, hjust = 0.5, face = 'bold'),
    axis.ticks = element_blank(),
    axis.title = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.key.size = unit(0.3, 'cm')
  )
)
```



## Replacing the Student List Business

The student list market is characterized by a teetering oligopoly controlled by two racially/socioeconomically problematic testing agencies and the looming takeover of corporate and private equity interests. Is this the best we can do? Considering research which suggests that student lists are surprisingly important for college access [@RN4739; @RN4752], regulating the student list business may be insufficient. Moreover, regulating existing products without introducing viable alternatives may cripple a mechanism that is vital for college access. Therefore, we sketch the contours of a "public-option" student list product that complements fundamental characteristics of the U.S. higher education system.

We conceive of U.S. higher education as a market-based system in which government takes a "steer don't row" approach, utilizing both subsidies and regulations to incentivize postsecondary education providers to achieve government goals. Drawing from principal-agent theory [@RN2673], federal and state governments (principals) want students to acquire human capital and contract with postsecondary institutions (agents) to deliver human capital [@RN3578]. State governments provide supply-side subsidies to public institutions in the form of state appropriations. The federal government (and most states) provide demand-side subsidies to students in the form of grants and loans, thereby creating a voucher system whereby public funding follows students to postsecondary institutions. Thus, postsecondary institutions (agents) have a financial incentive to provide access to students who receive financial aid from the federal/state government (principals). Student lists connect postsecondary institutions looking for students to prospective students looking for educational opportunities. At present, the student list business is oriented to the enrollment preferences of postsecondary institutions -- the customers paying oligopoly prices for names -- rather than equality of opportunity for students.

__The public option__. We propose a public-option student list product based on high school administrative data that is inspired by the example of national U.S. voter databases created by political parties [@RN4731]. The fundamental input to national voter databases are voter files created by state and local governments, which are essentially free public records. By contrast, the fundamental inputs for College Board and ACT student lists (e.g., contact information, academic achievement, college preferences) are proprietary. We argue that high school administrative data (e.g., contact information, GPA, course-taking by subject) provides the basis for a viable student list product. Increasingly, states have incorporated high school administrative data into statewide longitudinal data systems. The Arizona department of Education provides the contact information and high school rank of all AZ high school graduates for free to Arizona (public?) postsecondary institutions [CITE].

Therefore, we propose a student-list product developed by a consortium of states, based on data from statewide longitudinal data systems. Students and parents would have the opportunity to opt in or out. The "names" of students who opt in would be provided to eligible postsecondary institutions for free. The public-option student lists would include the following variables: contact information (mailing address, email, phone); high school code; expected graduation year; high school academic achievement (e.g., high school GPA, math course-taking, science course-taking); demographic information (e.g., gender, race, ethnicity); and limited set of survey responses about college preferences (e,g., intended major). 

Our analysis of the student list market suggests several product features that would enable the public-option to be a viable alternative to College Board/ACT products. First, the coverage of College Board and ACT student lists will likely wane as the test-optional admissions movement continues. The public-provision product becomes a viable alternative to the extent that it includes more states and a higher share of prospects from these states. Second, a core advantage of the College Board/ACT products is the ability to reach out to prospects early in the college search process (e.g., PSAT, PreACT). Therefore, the public-provision product should enable universities to obtain names of sophomores and juniors who have opted in, a product feature that requires timely data. Third, the students opting in to the public-provision product should be required to complete a very brief survey that provides information not recorded in high school administrative data (e.g., intended major, how far away from home student wants to be, preferences for public vs. private, etc.). Fourth, students/parents may want control over which institutions can obtain their contact information (e.g., in-state vs. out-of-state, institutional control).

If political and technical challenges can be overcome (described below), the public-option student list will improve college access and equality of opportunity in several ways. First, the segment targeting features of student lists (e.g., geodemographic filters, filtering by zip-code) that can lead to systematic exclusion of underrepresented students are a response to the high price College Board and ACT charge for names. By offering names for free, the public option eliminates the incentive for "efficient" name buys that target very particular prospects while excluding others. Free names do not prevent systematic exclusion in which prospects receive which recruiting interventions. Prior research finds that under-recruited student populations tend to be the most responsive to attention from universities [@RN4324]. Given that several recruiting interventions have negligible marginal cost (e.g., email, text), we argue that free names will result in more recruiting interventions directed at under-recruited populations. However, another likely consequence of free names is that students are deluged with emails and text messages.

A second benefit of the public-option is that students would not be excluded from name buys because they did not take a particular standardized assessment. Moving away from reliance on College Board and ACT databases for name buys will likely increase socioeconomic and racial equality in college access. Moreover, given the expected long-term decline in the number of SAT/ACT test-takers, creating the public-option student list may necessary to avoid a college access crisis. Finally, a third benefit of the public-option is that money previously spent on buying names can be allocated to other purposes. Furthermore, postsecondary institutions unable to afford names in the past (e.g., many community colleges) could engage in list-based recruiting. 

<!-- 
CUT TEXT

The responsible organization would provide student lists to "third party" universities. 

-->

__Consumer reporting agency designation__. The organization responsible for the public-option student list should be designated as a "consumer reporting agency" (CRA), thereby regulated by the Equal Credit Opportunity Act (ECOA), the Consumer Finance Protection Act (CFPA) and related laws. A CRA is defined as an entity that engages in "assembling or evaluating consumer credit information or other information on consumers for the purpose of furnishing consumer reports to third parties" [15 U.S.C \S 1681a ](https://www.law.cornell.edu/uscode/text/15/1681a). 

The rationale for the CRA designation is that the acquisition of student lists is the first step in the "enrollment funnel" process, which seeks to convert prospects into paying customers. Based on the information included in lists, universities decide which prospects to target and which recruiting interventions targeted prospects receive. Recruiting interventions are designed to solicit applications. Next, admitted applicants receive financial aid offers -- some combination of institutional grant aid and loan guidance -- designed to solicit enrollment and tuition payment. Thus, universities use student lists to identify potential customers who are systematically funneled towards taking on loans.

Recent actions by the National Student Clearinghouse suggest that organizations capable of running the public option student list may be unwilling to take on the CRA designation. The National Student Clearinghouse is a nonprofit organization that was founded to overcome data sharing problems that undermined that functioning of the federal financial aid system, problems that could not be solved by the federal government due to the prohibition of a national student unit record system. Amongst other services, the Clearinghouse provides verification services (e.g., degree verification, enrollment verification) which enable postsecondary institutions, high schools, loan servicers, employers, and other entities to share sensitive information about students. Clearly, the Clearinhouse possesses expertise relevant to the public option student list. However, in 2019 a group of students sued the Clearinghouse, arguing that the organization met the criteria of a consumer reporting agency and therefore should not have charged students \$30 to obtain a copy of their report [@pierce2020]. The Clearinghouse settled the lawsuit but stated that it "vigorously denies\ldots that it is a consumer reporting agency and that the FCRA…appl[ies] to it or its business practices" [CITE].

The lesson is that private-sector organizations with the technical capacity to run the public option student list do not want to be designated as consumer reporting agencies because of the additional regulatory requirements. However, the CRA designation is an essential element for a product that prioritizes consumer protection and equality of opportunity for students over other interests. Therefore, we argue that a government entity should assume responsibility for the public option student list.

__Political and technical challenges__. The public option faces political and technical barriers. The first political challenge is who pays to develop the system and maintain the system? Second, state laws restrict sharing statewide longitudinal education data. States with restrictive laws may be unable to participate in the public option. Third, state policymakers concerned about brain-drain may restrict out-of-state universities from obtaining the contact information of state residents. If several states adopt this approach, the public option becomes less viable as an alternative to College Board/ACT lists.

Technical challenges relate to creating a system that incorporates timely data -- contact information, transcript data, demographic information, survey responses -- from several states and from many districts and high schools within each state. With respect to the data generating process, state longitudinal education data systems incorporate data received from districts and from individual schools. Student information systems utilized by individual districts and schools vary tremendously in their capabilities, including their ability to share transcript data in a timely manner. 

High school transcript data are not standardized. Grading scales and course content differ across schools. A course named "algebra 1" may have very different content across schools, even for schools in the same state. In XXXX, the National Student Clearinghouse received funding to develop a system -- piloted in eight states -- that efficiently collected transcript data directly from schools [CITE]. However, efforts to analyze these data were frustrated because transcripts from different schools were fundamentally not comparable. Ironically, this challenge of evaluating the academic record of students from different high schools was a primary rationale for the creation of standardized college entrance exams [@RN2247]!

We are cautiously optimistic about overcoming these transcript comparison challenges for two reasons. First, the National Center for Education Statistics (NCES)
has completed national data collections of high school transcripts as part of the Longitudinal Studies Program since 1982 and as part of the National Assessment of Educational Progress (NAEP) since 1987 [@nces_hsts]. NCES could provide guidance about how standardizing transcript data for the public-option student list. Developing a standardized transcript data base does not ensure that courses with the same course-code have the same content across schools. Nevertheless, our second reason for optimism is that advances in data science make achieving apples-to-apples comparisons less important. Enrollment management professionals -- consulting firms and also sophisticated in-house operations -- typically have detailed data about each high schools and the postsecondary success of students from these schools. These data can be utilized to overcome concerns about comparing GPA or course-taking records across high schools.


# Recommendations for Universities and Admissions Professionals

PUBLIC UNIVERSITIES HAVE A RESPONSIBILITY TO EQUALITY OF OPPORTUNITY; THAT ISN'T THE SAME THING AS MEETING ENROLLMENT GOALS AND IT SHOULDN'T BE APPROACHED WITH THE SAME MINDSET OF "EFFICIENCY"

UNIVERSITIES

- LONG-TERM INVESTMENT IN THE IN-HOUSE CAPACITY OF YOUR ENROLLMENT OFFICE AND ADMISSIONS PROFESSIONALS



# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
