---
title: "The Student List Business"
  # "Deconstructing the student list business in higher education"
author: 
  - Ozan Jaquette
  - Karina Salazar
  - Patrica Martín
bibliography: ./../assets/bib/student_list_policy.bib # "./../../../Google Drive/student_list_brief/literature/student_list_policy.bib"
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../assets/bib/apa.csl
  # C:\Users\ozanj\Documents\student_list_policy\assets\bib
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 12pt
---

```{r, include = FALSE, eval = FALSE}
# filepath location of apa.csl file and location of .bib list of references
getwd()
list.files(path = file.path('.','..','assets','bib')) # apa.csl file
list.files(path = file.path('.','..','..','..','Google Drive','student_list_brief','literature')) # .bib reference list on Google Drive
```


```{r setup, include = FALSE}
library(knitr)
library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(scales)

knitr::opts_chunk$set(echo = F, message = F, warning = F)

knitr::knit_hooks$set(inline = function(x) {   if(!is.numeric(x)){     x   }else{    prettyNum(round(x,2), big.mark=",")    } })

theme_set(
  theme(
    text = element_text(size = 7),
    panel.background = element_blank(),
    plot.title = element_text(color = '#444444', size = 7, hjust = 0.5, face = 'bold'),
    axis.ticks = element_blank(),
    axis.title = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.key.size = unit(0.3, 'cm')
  )
)
```



## Replacing the Student List Business

The student list market is characterized by a teetering oligopoly controlled by two racially/socioeconomically problematic testing agencies and the looming takeover of corporate and private equity interests. Is this the best we can do? Considering research which suggests that student lists are surprisingly important for college access [@RN4739; @RN4752], regulating the student list business may be insufficient. Moreover, regulating existing products without introducing viable alternatives may cripple a mechanism that is vital for college access. Therefore, we sketch the contours of a publicly-provided student list product that complements fundamental characteristics of the U.S. higher education system.

We conceive of U.S. higher education as a market-based system in which government takes a "steer don't row" approach that combines subsidies and regulations to incentivize postsecondary education providers to achieve government goals. Drawing from principal-agent theory [@RN2673], federal and state governments (principals) want students to acquire human capital and contract with postsecondary institutions (agents) to deliver human capital [@RN3578]. State governments provide supply-side subsidies to public institutions in the form of state appropriations. The federal government (and most states) provide demand-side subsidies to students in the form of grants and loans, thereby creating a voucher system whereby public funding follows students to postsecondary institutions. Thus, postsecondary institutions (agents) have a financial incentive to provide access to students who receive financial aid from the federal/state government (principals). Student lists connect postsecondary institutions looking for students to prospective students looking for educational opportunities. At present, the student list business is oriented to the enrollment preferences of postsecondary institutions -- the customers paying oligopoly prices for names -- rather than equality of opportunity for students.

We propose a publicly-provided student list product based on high school administrative data that is inspired by the example of national U.S. voter databases created by political parties [@RN4731]. The fundamental input to national voter databases are voter files created by state and local governments, which are essentially free public records. By contrast, the fundamental inputs for student lists (e.g., contact information, academic achievement, college preferences) are proprietary. We argue that high school administrative data (e.g., contact information, GPA, course-taking by subject) provides the basis for a viable student list product. Increasingly, states have incorporated high school administrative data into statewide longitudinal data systems. Thus, we propose a student-list product developed by a consortium of states, based on data from statewide longitudinal data systems. Students and parents would have the opportunity to opt in or out. The "names" of students who opt in would be provided to eligible postsecondary institutions for free, eliminating the incentive for student list purchases that target specific prospects.

Our analysis of the student list market suggests several product features that would enable the publicly-provided product to be a viable alternative to College Board/ACT products. First, the coverage of College Board and ACT student lists will likely wane as the test-optional admissions movement continues. The public-provision product becomes a viable alternative to the extent that it includes more states and a higher share of prospects from these states. Second, a core advantage of the College Board/ACT products is the ability to reach out to prospects early in the college search process (e.g., PSAT, PreACT). Therefore, the public-provision product should enable universities to obtain names of sophomores and juniors who have opted in, a product feature that requires timely data. Third, the students opting in to the public-provision product should be required to complete a very brief survey that provides information not recorded in high school administrative data (e.g., intended major, how far away from home student wants to be, preferences for public vs. private, etc.). Fourth, students/parents may want control over which institutions can obtain their contact information (e.g., in-state vs. out-of-state, institutional control).

<!-- 
CUT TEXT

The responsible organization would provide student lists to "third party" universities. 

-->

We argue that the organization responsible for the publicly-provided product should be designated as a "consumer reporting agency" (CRA) defined as an entity that engages in "assembling or evaluating consumer credit information or other information on consumers for the purpose of furnishing consumer reports to third parties" [15 U.S.C \S 1681a ](https://www.law.cornell.edu/uscode/text/15/1681a). The rationale for the CRA designation is that the acquisition of student lists is the first step in the "enrollment funnel" process, which seeks to convert prospects into paying customers. Based on the information included in lists, universities decide which prospects to target. Targeted prospects receive recruiting interventions designed to solicit applications. Admitted applicants receive financial aid offers -- some combination of institutional grant aid and loan guidance -- designed to solicit enrollment and tuition payment. Thus, universities use student lists to identify potential customers who are systematically funneled towards taking on loans.

CRAs are regulated by the Equal Credit Opportunity Act (ECOA), the Consumer Finance Protection Act (CFPA) and related laws. In turn, the CRA designation raises the issue of what sort of organization is willing and able to be responsible for the publicly-provided product. 

For example, the non-profit organization National Student Clearinghouse provides verification services (e.g., degree verification, enrollment verification) that enable postsecondary institutions, loan servicers, employers, and other entities to share confidential information about students. The Clearinghouse was founded to overcome data sharing challenges for the federal financial aid system that could not be solved by the federal government because of the prohibition of a national student unit record system. The Clearinghouse posseses expertise relevant to the creation of a publicly-provided list product. However, in 2019 a group of students sued the Clearinghouse, arguing that the organization met the criteria of a consumer reporting agency and therefore should not have charged students \$30 to obtain a copy of their report [@pierce2020]. The Clearinghouse settled the lawsuit but stated that it "vigorously denies…that it is a consumer reporting agency and that the FCRA…appl[ies] to it or its business practices" [CITE]. 

The lesson, we believe, is that private-sector organizations with the technical capacity to run the publicly-provided product do not want to be designated as consumer reporting agencies because of the additional regulatory requirements. However, CRA designation is essential for maintaining a publicly-provided product that protects students. Therefore, we argue that a government entity should assume responsibility for the publicly-provided product. One potential approach is a partnership between states, with coordinating support from the federal government.


However, in 2019 a group of students filed 


For example, the National Student Clearinghouse is a non-profit organization that was founded to facilitate data sh


Thousands of companies and organizations nationwide depend on the Clearinghouse to take the guesswork out of confirming academic credentials. We also offer industry-based and custom solutions that leverage our unique education data resources.


The National Student Clearinghouse 


 raising the issue of what sort of organization is willing and able to be responsible for the publicly-provided product. 

cite this guy @pierce2020


Two fundamentally important decisions are what sort of organization would be responsible for the product and how would this organization be regulated. We argue that the responsible organization should be designated as a "consumer reporting agency," defined as an entity that engages in "assembling or evaluating consumer credit information or other information on consumers for the purpose of furnishing consumer reports to third parties" [15 U.S.C \S 1681a ](https://www.law.cornell.edu/uscode/text/15/1681a). 


# Recommendations for Universities and Admissions Professionals

PUBLIC UNIVERSITIES HAVE A RESPONSIBILITY TO EQUALITY OF OPPORTUNITY; THAT ISN'T THE SAME THING AS MEETING ENROLLMENT GOALS AND IT SHOULDN'T BE APPROACHED WITH THE SAME MINDSET OF "EFFICIENCY"

UNIVERSITIES

- LONG-TERM INVESTMENT IN THE IN-HOUSE CAPACITY OF YOUR ENROLLMENT OFFICE AND ADMISSIONS PROFESSIONALS



# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
