---
title: "The Student List Business in Higher Education"
author: 
  - Ozan Jaquette
  - AUTHOR TWO
  - Patrica Martin
bibliography: "./../../../Google Drive/student_list_brief/literature/student_list_policy.bib"
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../assets/bib/apa.csl
  # C:\Users\ozanj\Documents\student_list_policy\assets\bib
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 11pt
---

```{r, include = FALSE, eval = FALSE}
# filepath location of apa.csl file and location of .bib list of references
getwd()
list.files(path = file.path('.','..','assets','bib')) # apa.csl file
list.files(path = file.path('.','..','..','..','Google Drive','student_list_brief','literature')) # .bib reference list on Google Drive
```


```{r setup, include = FALSE}
library(knitr)
library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
library(tidyverse)
#library(gridExtra)
#library(scales)

knitr::opts_chunk$set(echo = F, message = F)

knitr::knit_hooks$set(inline = function(x) {   if(!is.numeric(x)){     x   }else{    prettyNum(round(x,2), big.mark=",")    } })
```

  The student list business is broadly defined as a business focused on the selling and purchasing of student information. A 2018 report by the Fordham Law School Center on Law and Information Policy titled, _Transparency and the Marketplace for Student Data_ details the purchasing of student lists in the commerical marketplace and the need for student data privacy laws [@russell2018transparency]. Student data is appealing to businesses who want to make a profit from marketing their products to a younger demographic [@russell2018transparency]. Some businesses include clothing retailers, sports and recreation facilities, and to our concern, colleges and universities [^1]. In their study, @russell2018transparency found at least 14 data brokers-- that is businesses who primarily focus on collecting, aggregating, analyzing, and sharing consumer information-- sent marketing solicitations by mail or email to 10 high school students [^2]. While the authors note that it was relatively easy to identify the data brokers sending marketing solicitations to students, it was a lot more challenging to pinpoint where the data brokers obtained students' information [@russell2018transparency].  
  Through their data collection, @russell2018transparency found that data brokers were able to gather information on students from some of the following sources: student surveys and questionnaires, affiliations and/or coordination with other brokers, and other unknown and less transparent sources [@russell2018transparency]. For example, @russell2018transparency found that the Educational Research Center of America (ERCA) collects student information through surveys they administer at schools. Furthermore, ERCA also shares information with ASl Marketing LLC, a for-profit company that engages in selling student data to commercial entities that market their products and services to that specific demographic [@russell2018transparency]. Businesses interested in marketing their products to a younger student demographic rely on vendors or data brokers to provide them with student data-- colleges and universities also rely on this information.  
  Recent news articles have shed light on the practice of vendors such as College Board and ACT collecting student information and selling it to buyers (e.g. universities, marketing companies)[@singer2018-studata; @belkin2019-studata]. A New York Times article highlighted the ways that students are targeted for educational and noneducational marketing by companies like MyCollegeOptions-- who was owned by the National Research Center for College and University Admissions (NRCCUA) which is now owned by ACT [@singer2018-studata]. @singer2018-studata also notes how Scholarships.com collects information from students and shared that information with American Student Marketing, a data broker of the "student data collected by Scholarships.com, offering it to marketers of student loans, credit cares and smartphones." Some concerns about the use of student data is in the ways that vendors or "survey services may categorize students in ways that could expose them to predatory marketing or exclude them from important opportunities."  
  The process of categorizing students... segmentation.  
  Vendors are particularly successful at homing in on target student populations through a sophisticated system known as geodemographics. Geodemographics is the process of combining enormous public and private data sources to gather information on consumer behavior, using Geographic Information Systems (GIS) to spatially analyze markets, and segmentation schemes to classify consumers into clusters of ideal prospects (Goss, 1995). These systems have been used in politics, real estate, financial services, marketing, and more to inform decision-making and increase efficiency (Goss, 1995).  [Maybe include an example of this historically in practice]. Ultimately the goal of geodemographics is to enable marketers to “... predict behavioral responses of consumers based on statistical models of identity and residential location” (Goss, 1995, p. 171). While it may come at no surprise, geodemographics has been used in higher education marketing to target specific populations of students [@college-board2011](CITE news articles on this).  
  Geodemographics is tied to Jonathan Robbin who is credited with pioneering a computer-based geodemographic system in the early 1970s, which eventually led to the creation of PRIZM (Potential Rating Index for ZIP Markets) [@burrows2006geodemographics; @goss1995we]. Using his training in sociology, demography, and statistics, Robbin transformed “... the iterative factorial ecologies of positivist urban social science into a powerful marketing tool” [@burrows2006geodemographics; @goss1995we, p. 173]. In a similar fashion, Richard Webber, a social scientist in the UK began to develop a geodemographic classification system called ACORN (A Classification of Residential Neighborhoods) and later (MOSAIC) to profile customers using their postcodes [@burrows2006geodemographics]. These software systems have reached other countries including Canada and Australia. Today, Claritas, the company founded by Robbin, and owners of PRIZM, operate in the US and provide their services to big brand companies including the likes of McDonald's, The Home Depot, Univision, Salesforce, and more [CITE]. Claritas promote their proprietary “Claritas Identity Graph” which they claim “reaches nearly 100% of US consumer households and ties together over 5 billion data points monthly to produce the highest-def portrait of each customer and prospect” [CITE]. Claritas data is derived from gathering over 10,000 predictive demographic and behavioral indicators using demographic information (e.g. age, income, home ownership, wealth) about consumers, multicultural behaviors (e.g. ethnicity, language at home, acculturation), lifestyle behaviors (e.g. shopping, media, insurance) and purchase behaviors (e.g. device ownership, service provider, restaurant & retail preferences)[CITE].  
  Transition to discuss in more detail how the actual process of geodemographics works. Geodemographics is premised on ‘the fundamental sociological truism that “birds of a feather flock together”’ meaning those who share similar demographics, values, and characteristics tend to share proximity in social and physical space [as cited in @goss1995we, p. 172; @burrows2006geodemographics; @college-board2011]. Through this idea, the practice of geodemographics is enacted through obtaining detailed aggregated information on individuals and classifying them based on demographic and psychographic characteristics [@goss1995we]. The process broadly requires three components: massive electronic data of individuals, Geographic Information Systems (GIS) or related technology to spatially represent individuals, and segmentation schemes to identify prospects through factor and cluster analysis [@goss1995we].  As mentioned previously, geodemographics has been used for various purposes, particularly for targeted marketing in retail and businesses, but this practice of purchasing consumer data for marketing purposes is no stranger to higher education [@college-board2011].  
  In a 2011 Segment Analysis Service report, the College Board detailed their geodemographic tagging service, which unlike mainstream geodemographic systems, is targeted entirely at college-bound students [@college-board2011]. The College Board gathers this information through publicly available US Census data and proprietary data obtained [FOR EXAMPLE] from students who register for tests administered by the College Board (e.g. PSAT/SAT, AP) [@college-board2011]. Once this information is obtained from students, other data elements such as test scores and test score sending patterns are attached to each student record and then grouped in two ways: at the high school level and neighborhood level [@college-board2011]. This process culminates in unique behavioral profiles for approximately 33,000+ high schools and 44,000+ educational neighborhoods in the United States that are then each complemented with 150 attributes (i.e. based on college information, educational and occupational attainment, population demographics, etc.) [@college-board2011]. This massive amount of data is then reduced to a descriptive set of 40 unique high-school factors and 38 educational neighborhood factors to be weighted and representative of the larger set of individual attributes for each group [@college-board2011]. Lastly, the unique factors from the previous step are used to create 29 unique high school types and 33 unique neighborhood types, also referred to as clusters [@college-board2011]. This process of segmentation allows for the clustering or grouping of members who are “more like each other than they are like members of other clusters… Ultimately, each student with a home address can be mapped to one of the neighborhood clusters, and the student’s high school can be mapped to one of the high-school clusters, providing detailed insights into his/her likely characteristics and college-bound behaviors.” [@college-board2011, p. 4]. This report was written almost a decade ago and while the details of the process may have changed, one could imagine that the College Board and other student list vendors continue to use geodemographics to profit from student data.  
  Geodemographics and related targeted marketing systems are not devoid of using race, socioeconomic status, and other demographic characteristics to target prospects. The practice of identifying prospects based on their home address to capture homogenous communities fails to acknowledge racial segregation rooted in the system of slavery in the U.S. [@rothstein2017color]. Racial segregation is not a cause of unintentional happenstance, but rather a result of systematic policies-- private prejudice, white flight, real estate steering, bank redlining [@rothstein2017color]. Therefore it is no coincidence in the ways neighborhoods are spatially and socially organized so that poor neighborhoods tend to be home to predominantly black, indigenous, and people of color and wealthier neighborhoods (suburbs) tend to be home to white, more affluent communities [@rothstein2017color]. The socio-spatial organization of neighborhoods is a consequence of the racial domination and economic exploitation of Black and Native American people or what Professor Cheryl Harris calls whiteness as property [@harris1993whiteness].  
  Professor Harris discusses how whiteness and property collectively serve as systems of domination of Black and Native American people through their mutual process of exclusion [@harris1993whiteness]. Whiteness as property has taken shape throughout history beginning with slavery and the seizure of indigenous lands to the modern legal protection of white privilege [@harris1993whiteness].  @harris1993whiteness argues that slavery as a legal institution upheld the right to treat enslaved people as property and through this system, racial identity was conflated with slavery. She writes, “... it became crucial to be 'white,' to be identified as white, to have the property of being white. Whiteness was the characteristic, the attribute, the property of free human beings.” [@harris1993whiteness, p. 1721]. Furthermore, the seizure of indigenous lands was further perpetuated and justified by a racial and cultural “otherness” [@harris1993whiteness, p. 1721]. In other words, the right to property was conferred to white people alone [@harris1993whiteness]. While property in the traditional sense represents everything of value that a person has a right to, @harris1993whiteness maintains that the modern concept of property is reinforced and protected, albeit implicitly and explicitly, by the law.  
  Through the creation of property “rights”, the law determines whose interests are protected and how these rights are enforced [@harris1993whiteness]. The resulting inequalities, Harris argues, influence the structuring of social relations and the construction of "'whiteness' as an objective fact” [@harris1993whiteness, p. 1731]. This upholding of whiteness in the law also has implications for the property functions of whiteness and those who are afforded these privileges [@harris1993whiteness]. @harris1993whiteness outlines the four property functions of whiteness to include: 1) the rights of disposition, 2) rights to use and enjoyment, 3) reputation and status property, and 4) the absolute right to exclude. In describing how a new form of whiteness as property emerged, Harris details the _Brown v. Board of Education_ Supreme Court case [@harris1993whiteness].  
  The _Brown v. Board of Education_ court case was a watershed moment in U.S. history in a decision where the Supreme Court reversed the ruling in _Plessy v. Ferguson_ of  “separate but equal”-- abolishing de jure segregation [@harris1993whiteness]. @harris1993whiteness argues that the Brown case legally prohibited racial segregation, but it did not address subtle forms of racial segregation in race-based privileging of white people. @harris1993whiteness argues that the Brown case “... dismantled an old form of whiteness as property while simultaneously permitting its reemergence in a more subtle form. White privilege accorded as a legal right was rejected, but de facto white privilege not mandated by law remained unaddressed.” (p. 1753). This unchecked white privilege in society has contributed to higher rates of school segregation today than there were forty years ago [@rothstein2017color]. The ruling in the Brown case therefore has led to the preservation of unequal resources in public education and subsequent cases (e.g. _Milliken v. Bradley_) failing to acknowledge how residential segregation and school segregation work in tandem to marginalize communities of color [@harris1993whiteness]. The right to property manifests itself in both the tangible ownership of land and the intangible ownership of knowledge [@harris1993whiteness; @salazar2019wealth].  
  @ladson2006toward delve into the intersections of race and property and its impact on educational inequities. Informed by Harris's analysis of the property functions of whiteness outlined above, they note that whiteness as property preserves the unequal distribution of both physical (tangible) and intellectual (intangible) resources of schools serving Black students [@harris1993whiteness; @ladson2006toward; @salazar2019wealth]. The property differences in schools materialize in different ways such as in the quality and quantity of the curriculum-- a form of “intellectual property” [@ladson2006toward, p. 54]. Intellectual property refers to both access to physical resources (e.g. computers, state-of-the-art technologies, certified teachers and staff) and the curriculum (e.g. standards of what students should learn and be able to do) [@ladson2006toward]. Another representation of whiteness as property is in the reputation and status of schools serving predominantly Black, Indigenous, and Students of Color and their associations with being non-white or “urban” [@salazar2019wealth]. Schools that are associated with these characteristics are deemed "'disadvantaged'" and "'underprepared'" and Students of Color are cast in a deficit lens [@ladson2006toward; @salazar2019wealth, p. 30]. The associations historically between race and space (e.g. neighborhoods, high schools) have real life consequences for how student lists purchased on the basis of students' zip codes and/or other sophisticated geodemographic systems have on college access.  
  Geodemographics is grounded in the social networks principle of homophily-- that is, people's personal networks are more alike [@college-board2011; @mcpherson2001birds]. The scholarship on the intersections of race and property demonstrates that the socio-spatial organization of neighborhoods is an outcome of racism, white supremacy, de jure and de facto segregation-- or the actualization of whiteness as property [@harris1993whiteness; @rothstein2017color]. Therefore, geodemographics is a sophisticated system that is used with the intention of reaching specific groups of people. The College Board’s segmentation analysis technology identifies particular neighborhoods and high schools-- both of which have implications for how universities associate characteristics given about these cluster areas [@ladson2006toward; @salazar2019wealth]. Schools that are deemed “urban” may be overlooked by colleges and universities because they are associated with negative stereotypes [@ladson2006toward; @salazar2019wealth]. Literature on marketing and recruiting in public and for-profit institutions further substantiates the intentional marketing practices that prioritize particular students while ignoring other students [@dache2018dangling; @han2019recruiting; @cottom2017lower; @salazar2019wealth].  
  Research on the off-campus recruiting of 15 public research universities has found that universities tend to prioritize their recruiting efforts in out-of-state, wealthy, and/or high schools serving predominantly white students [@han2019recruiting]. @salazar2019wealth extends this research by providing a critical analysis of recruiting visits made by public research universities in the Los Angeles and Dallas metropolitan areas. Her work draws from critical geography [@agnew2011space] and whiteness as property [@harris1993whiteness] in exploring the relationship between the spatial distributions of Communities of Color and their proximity to resources (both tangible and intangible) [@salazar2019wealth]. @salazar2019wealth finds that universities perpetuate what @harris1993whiteness calls a modern view of whiteness where universities recruit in wealthy, predominantly white communities, while ignoring communities that have a status or reputation of being non-white. This modern function of whiteness in off-campus recruiting of public research universities serves to exclude Black, Indigenous, and People of Color from enrolling in these institutions. Although the research on the marketing and recruiting practices of public research universities is fairly new, similar tactics have been used for decades by for-profit institutions, albeit to recruit different student populations [@dache2018dangling; @cottom2017lower].  
  The predatory recruitment practices of for-profit institutions is well documented in the literature [@deming2012profit; @kutz2010profit; @senatehealth]. @cottom2017lower provides an account of the conditions that led to the new economy and the labor market failures that became profitable for for-profit institutions. The declining investment in social insurance policies (e.g. health insurance and pensions) also led to a declining investment in public higher education, which in turn left students and families vulnerable to accrue more student debt-- or what @cottom2017lower refers to as “Lower Ed” (p. 16). Lower Ed is predicated on deteriorating investment in public education and poor labor market conditions, which the for-profit industry is able to thrive off of [@cottom2017lower]. For-profit institutions spend a whole slew of resources on marketing and recruiting to those vulnerable to these social and economic realities-- in particular Black women [@cottom2017lower]. Research shows that the marketing and location of for-profit institutions is not a coincidence and on the contrary, for-profit institutions tend to be located in urban areas that are home to communities of color [@dache2018dangling]. The proximity of these institutions to communities of color allows for-profit institutions to recruit students using factors of race, gender, and class [@dache2018dangling]. The literature on the recruitment practices of public research universities and for-profit institutions demonstrates the ways in which race and geographical space are pervasive in the US and continue to be used to leverage marketing and recruiting goals. The student list business is one small, yet critical tactic used by institutions for the targeting marketing and recruitment of students. 

[^1]: Recommendations from https://www.infousa.com/lists/student-mailing-lists/   
[^2]: For full report see https://www.fordham.edu/info/23830/research/10517/transparency_and_the_marketplace_for_student_data

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
